Computing {\em the least commitment} action model from observations of plan executions

### 2018-11-26 doc/ijcai-19-conformant/conformant-learning.tex
added first definition for least commitment action model

A solution to a conformant planning problem is an action sequence that maps each possible initial state into a goal state. More precisely an action sequence $\pi=\tup{a_1, \ldots, a_n}$ is a {\em conformant plan} for $P_c$ iff for each possible {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0$ is a valuation that satisfies $\Upsilon$ and it holds that the trajectory reaches $G \subseteq s_n$.

@@

\subsection{The hypothesis space}
\strips\ action schemata provide a compact representation for specifying classical planning models. {\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.

@@

For instance, the {\small\tt stack(?v1,?v2)} and {\small\tt unsstack(?v1,?v2)} schemata from a four operator {\em blocksworld}~\cite{slaney2001blocks} are {\em comparable} while {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} are not. Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schemata $\xi'\in\mathcal{M'}$ and vice versa.

@@

\item {\em Domain-specific constraints}. One can introduce domain-specific knowledge to constrain further the space of possible schemata. For instance, in a {\em blocksworld} one can claim that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in any of these lists because of the semantic of the {\tt\small on} predicate that codes which block is in top of other block. 
\end{enumerate}

@@

\begin{enumerate}
\item[3.] {\em Observation constraints}. An observation of the execution of a plan further constraints the space of possible action schemata. This \emph{semantic knowledge} included in the observations introduce a third type of constraints, that we will call {\em observation constraints}, and that can be added to the set $\mathcal{C}$.
\end{enumerate}

\subsection{The learning task}

>>>>

A solution to a conformant planning problem is an action sequence that maps each possible initial state into a goal state. More precisely an action sequence $\pi=\tup{a_1, \ldots, a_n}$ is a {\em conformant plan} for $P_c$ iff for each possible {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0$ is a valuation of the fluents in $F$ that satisfies $\Upsilon$ and it holds that the trajectory reaches $G \subseteq s_n$.

@@

\subsection{The hypothesis space}
\strips\ action schemata provide a compact representation specifying action models. {\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.

@@

For instance, the {\small\tt stack(?v1,?v2)} and {\small\tt unstack(?v1,?v2)} schemata from a four operator {\em blocksworld}~\cite{slaney2001blocks} are {\em comparable} while {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} are not. Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function that maps every action schema $\xi\in\mathcal{M}$ to a comparable schemata $\xi'\in\mathcal{M'}$ and vice versa.

@@

\item {\em Domain-specific constraints}. One can introduce domain-specific knowledge to constrain further the space of possible schemata. For instance, in the {\em blocksworld} one can claim that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of an action schema $\xi$ because of the semantic of the {\tt\small on} predicate (i.e. only one block can be on top of another block). 
\end{enumerate}

@@

\begin{enumerate}
\item[3.] {\em Observation constraints}. An observation of the execution of a plan further constraints the space of possible action schemata. This \emph{semantic knowledge} included in the observations introduce a third type of constraints and that can also be added to the set $\mathcal{C}$.
\end{enumerate}

\subsection{The least commitment action model}
Now we are ready to formalize the learning task that we are addressing in this paper. The task of computing {\em the least commitment} action model is defined as a tuple $\tup{P,M,\mathcal{O}}$ where:

\begin{itemize}
\item $P=\tup{F,A[\cdot],I,G}$ is a planning problem where $A[\cdot]$ is a set of actions. For each $a\in A[\cdot]$, the semantics of $a$ is unknown; i.e. the functions $\rho$ and/or $\theta$ of $a$ are undefined.
\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that solves the planning problem $P$.
\item $M$ is the set of different planning models for the actions in $A[\cdot]$. A model $\mathcal{M}\in M$ defines the semantics of every action in $A[\cdot]$. Planning models differ in the $\tup{\rho,\theta}$ functions of the actions but they all use the same set of state variables $F$.
\end{itemize}

The {\em solution} to this task is {\em the least commitment} action model which is defined as the minimum subset of models $M^*\in M$ that is compliant with the input observation $\mathcal{O}(\tau)$.
